Hyperbolicity of Augmented Links in the Thickened Torus

### 2021-09-01 main.tex
more figures
putting the figures back in context
diagram for bow-tie-modifications
diagram for crossing edges

@@

@@

\begin{tikzpicture}
\node[emptynode] (a) at (-0.9,-0.3) {};
\node[dotnode] (b) at (-0.3,-0.1) {};
\node[dotnode] (c) at (0.3,0.1) {};
\node[emptynode] (d) at (0.9,0.3) {};
%% two strands going out back
\draw (b) -- +(150:1cm);
\draw (c) -- +(150:1cm);
%% the ellipse; draw top arc, then fill, then bottom arc
\draw[thin_overline={1}] (a) -- (b) -- (c) -- (d)
	.. controls +(90:0.8cm) and +(90:0.8cm) .. (a);
\draw[fill, opacity=0.2] (a) -- (d)
	.. controls +(90:0.8cm) and +(90:0.8cm) .. (a);
\draw (a) -- (d)
	.. controls +(-90:0.8cm) and +(-90:0.8cm) .. (a);
%% two strands coming out front with twist
%\draw (b) -- +(-30:2cm); %\draw (c) -- +(-30:2cm);
\draw[thin_overline={1}] (c)
	.. controls +(-30:0.5cm) and +(150:0.5cm) .. (0.9,-0.8);
\draw[thin_overline={1}] (b)
	.. controls +(-30:0.5cm) and +(150:0.5cm) .. (1.5,-0.6);
\end{tikzpicture}

\comment{

@@

\begin{define}
Recall that a \emph{twist region}
in a link diagram is a maximal sequence of vertices
such that consecutive vertices are two vertices of a bigon face,
and consecutive bigons meet only at a vertex (not an edge).
We say that the \emph{direction} of a twist region
is the pair of faces (possibly the same face)
that are across from the bigons at the end;
if a twist region consists of only one vertex,
a direction is a pair of opposite faces (again, possibly the same face)
meeting at the crossing.
%%\footnote{There is an ambiguity of ``maximality'' in this definition:
%%for example, if a vertex of a link diagram meets exactly one bigon,
%%it can be considered as part of a twist region of length 2
%%(given by that bigon),
%%or it can be considered as part of a twist region of length 1,
%%as there are no bigons in the other ``direction''.
%%It should be clear which ``direction'' a twist region is going
%%when we deal with augmentations later.
%%}

@@

\begin{definition}
Let $G$ be a graph on the torus.
Let $v$ be a vertex and $e,e'$ be distinct edges that meet $v$.
A \emph{left (resp. right) bow-tie modification to $v,e,e'$}
is the process of removing $v,e,e'$ and adding in a ``bow-tie''
as in Figure \ref{TODO} (a) (resp. (b)).
We call the TODO long edges, the TODO short edges,
and TODO diagonal edges.
TODO in figure caption, left bow-tie modification so-called because
the triangular face appears to the left of the long edges;
also note that left/right conincide with Figure \ref{f:right_left_aug}
\label{d:diagram-bowtie}
\end{definition}

@@

\begin{figure}
\includegraphics[width=5cm]{more_pictures/right_left_aug.png}
\caption{R: right augmentation, L: left augmentation}
\label{f:right_left_aug}
\end{figure}

@@

We call the resulting graph the \emph{top bow-tie graph of $L$},
denoted by $\Gamma_T(L)$.
If we had oriented the edges of $D'$ the other way,
and subsequently performed the same operations,
we obtain another graph,
which we call the \emph{bottom bow-tie graph of $L$},
denoted by $\Gamma_B(L)$.
\hfill $\triangle$
\end{definition}

@@

We will begin by assuming that $L$ has no half twists.
Let $L = K \cup C$, with $C$ being the collection of crossing circles.
%Arrange the link diagram of $L$ in the following way:
Arrange $L$ in the following way:
place the circle components in $C$ perpendicular to
the projection plane $\torus \times \{0\}$,
and leave the remaining part of the link $K \subseteq L$
lying in the projection plane (except at crossings of $K$).
Thus, the projection of $L$ onto the projection plane
will be a diagram $D(K)$ of $K$
together with line segments corresponding to crossing circles.
(We demand that $D(K)$ be alternating and twist-reduced.)
As discussed before, we may remove full twists (i.e. pairs of bigons)
from the twist regions of $K$ that are augmented in $L$,
and since $L$ is assumed to have no half twists,
the two edges of $D(K)$ that go through a crossing circle
do not meet (see Figure \ref{fig:Augmentations} C).
From now on, we will take $K$ to be this modified link
(with full twists removed).

We now place a \emph{crossing edge} at each crossing of $K$,
connecting the top and bottom strands at the crossing
(see Figure \ref{fig:crossingArc} (a)).
%so that for each crossing edge,
%one end of the edge lies on a bottom strand while the other
%end lies on a top strand as in Figure \ref{fig:crossingArc} (a).

\begin{figure} 
\centering 
\begin{tabular}{cc}
\includegraphics[width=4cm]{crossingArc.png}
& \includegraphics[width=5cm]{crossingPush.png}
\\
(a)
& (b)
\end{tabular}
\caption{(a) The black strands are
part of the link and the red strand is the crossing edge.
(b) The blue and
red edges represent the split crossing edges and the shaded half disk is bounded
by the crossing circle} 
\label{fig:crossingArc} 
\end{figure}

Next we construct the graph that will define the ``top'' torihedron.
We view the link from the point at infinity from the top end
($\torus \times \{1\}$) of the thickened torus.
At each crossing of $K$,
push the top strand towards the bottom strand,
splitting the crossing edge into two
identical edges and spreading them apart
as in Figure \ref{fig:crossingArc} (b).
%We push the link
%components to infinity and stretch the crossing edge so that we have flattened
%the link onto $\torus \times \{0\}$ except for the crossing circles,
%which remains perpendicular to the projection plane.

@@

@@

@@

There exists a non-degenerate
%angle-splitting of $(\Gamma,\theta_\bullet)$
angle-splitting of $(\Gamma,f^*)$
if and only if the following
``cocycle condition'' is satisfied:

@@

Recall that $\Gamma_T(L),\Gamma_B(L)$ are
obtained by bow-tie modifications of the diagram $D'$
of a link $K'$
(see \defref{d:tori-decomp-graph}).
Assign to each edge $e$ of $D'$ the angle $\theta_e = \pi/2$
(so that $\theta_e^* = \pi/2$ too).
Using the fact that $K'$ is weakly prime
(which easily follows from $K$ being weakly prime),
it is not hard to see that the ``cocycle condition''
of \thmref{t:bs-thm4} is satisfied by this assignment.
Thus, there exists a non-degenerate angle-splitting
$\vphi_\bullet$ of $(D',\theta_\bullet^*)$.
%Recall from the proof of \prpref{p:tori_decomp}
%that the two graphs $\Gamma_T(L),\Gamma_B(L)$
%can be obtained by modifying the link diagram of $K$:
%we take $K'$ to be the link that agrees with $K$ everywhere
%except each augmented twist region has only one crossing,
%and $\Gamma_T(L),\Gamma_B(L)$ are

@@

It is straightforward to check that
upon gluing the top and bottom torihedra,
the sum of interior dihedral angles $\theta^*$
around each edge is $2\pi$;
crossing edges have $\theta^* = \pi/2$,
and appear four times, twice in each torihedron,
while for bow-tie edges,
simply check for half-twist and non-half-twist cases
separately.

Now we have a decomposition of the two torihedra into
degenerate base-angled pyramids
(recall that that means that some of the
interior dihedral angles $\theta^*$ are 0),
but we need the pyramids to be non-degenerate,
so we need to modify the graph on the torihedra
and the angle assignments to make all $\theta^*$ nonzero.

@@

\begin{figure}
\includegraphics[width=5cm]{more_pictures/one_edge.png}
%\includegraphics[width=5cm]{more_pictures/two_edge.png}
\includegraphics[width=5cm]{more_pictures/two_edge_many.png}
\caption{TODO change this. Red edge indicates an added edge to $\Gamma_T(L)$;
the +/- labels indicate whether to
increase/decrease $\theta^*$ and $\vphi$.}
\label{f:adding_edges}
\end{figure}

@@

Now fix some small $\veps > 0$.
Let $\vphi_{\vec{e}}' = \vphi_{\vec{e}} + x \cdot \veps$,
where $x$ is the label on $\vec{e}$ (0 if unlabeled).
Let $\theta_e'^* = \vphi_{\vec{e}}' + \vphi_{\cev{e}}'$.
It is easy to check that the sum of $\theta'*$
around each vertex is the same as for $\theta^*$,
so it defines an angled torihedron $(\sT_T',\theta'^*)$.
It is also easy to check that for each face of $\Gamma_T'$,
the $+/-$ labels cancel out,
so $\vphi'$ is an angle-splitting of it.
Furthermore, all shorts edges
(which are the only edges with $\vphi_{\vec{e}} = 0$)
have a $+1$ label on each side,
so $\vphi'$ is non-degenerate.

@@

It is easy to see that, under the gluing of
$\sT_T(L)$ to $\sT_B(L)$,
the new edges added to $\Gamma_T(L)$
are glued to the new edges added to $\Gamma_B(L)$.
As noted before, upon gluing $\sT_T(L)$ to $\sT_B(L)$
the sum of exterior dihedral angles $\theta^*$ around each edge
is $2\pi$.
This clearly remains true after adding the new edges
(they're label $\pi$ on each torihedron).
It is also eashy to check that the $+/-$ labels get canceled out.
Thus, upon gluing $\sT_T'$ to $\sT_B'$,
the sum of new exterior dihedral angles $\theta'^*$ around each edge
is $2\pi$.

@@

}

@@

%crossing edge, Menasco thing
%\label{fig:crossingArc}

>>>>

%% colors
\definecolor{light-gray}{gray}{0.9}
\definecolor{med-gray}{gray}{0.6}

@@

%%% \comment{

@@

@@

\begin{define}
Recall that a \emph{twist region}
in a link diagram is a maximal sequence of vertices
such that consecutive vertices are two vertices of a bigon face,
and consecutive bigons meet only at a vertex (not an edge).
We say that the \emph{direction} of a twist region
is the pair of faces (possibly the same face)
that are across from the bigons at the end;
if a twist region consists of only one vertex,
a direction is a pair of opposite faces (again, possibly the same face)
meeting at the crossing.
We say that a twist region consisting of only one vertex
is \emph{trivial}.
%%\footnote{There is an ambiguity of ``maximality'' in this definition:
%%for example, if a vertex of a link diagram meets exactly one bigon,
%%it can be considered as part of a twist region of length 2
%%(given by that bigon),
%%or it can be considered as part of a twist region of length 1,
%%as there are no bigons in the other ``direction''.
%%It should be clear which ``direction'' a twist region is going
%%when we deal with augmentations later.
%%}

@@

\begin{definition}
Let $G$ be a graph on the torus.
Let $v$ be a vertex and $e,e'$ be distinct edges that meet $v$.
A \emph{left (resp. right) bow-tie modification to $v,e,e'$}
is the process of removing $v,e,e'$ and adding in a ``bow-tie''
as in Figure \ref{TODO} (a) (resp. (b)).
We call the TODO \emph{long} edges, the TODO \emph{short} edges,
and TODO \emph{diagonal} edges,
and we call the new triangular faces the \emph{bow-tie faces}.
TODO in figure caption, left bow-tie modification so-called because
the triangular face appears to the left of the long edges;
also note that left/right conincide with Figure \ref{f:right_left_aug}

\begin{figure}[ht]
\input{bow-tie-modification.tikz}
\caption{Bow-tie modifications}
\label{f:bow-tie-modification}
\end{figure}

\label{d:diagram-bowtie}
\end{definition}

@@

@@

We call the resulting graph the \emph{top bow-tie graph of $L$},
denoted by $\Gamma_T(L)$.
If we had oriented the edges of $D'$ the other way,
and subsequently performed the same operations,
we obtain another graph,
which we call the \emph{bottom bow-tie graph of $L$},
denoted by $\Gamma_B(L)$.
%\hfill $\triangle$
\end{definition}

\begin{figure}[ht]
\input{torihedra-graphs.tikz}
\caption{Constructing the top and bottom bow-tie graphs,
using bow-tie modifications}
\label{f:tori-decomp-graph}
\end{figure}

@@

We will begin by assuming that $L$ has no half twists.
Let $L = K \cup C$, with $C$ being the collection of crossing circles.
%Arrange the link diagram of $L$ in the following way:
Arrange $L$ in the following way:
place the circle components in $C$ perpendicular to
the projection plane $\torus \times \{0\}$,
and leave the remaining part of the link $K \subseteq L$
lying in the projection plane (except at crossings of $K$).
Thus, the projection of $L$ onto the projection plane
will be a diagram $D(K)$ of $K$
together with line segments corresponding to crossing circles.
%(We demand that $D(K)$ be alternating and twist-reduced.)
As discussed before, we may remove full twists (i.e. pairs of bigons)
from the twist regions of $K$ that are augmented in $L$,
and since $L$ is assumed to have no half twists,
the two edges of $D(K)$ that go through a crossing circle
do not meet (see Figure \ref{fig:Augmentations} C).
From now on, we will take $K$ to be this modified link
(with full twists removed).

We now place a \emph{crossing edge} at each crossing of $K$,
connecting the top and bottom strands at the crossing
(see \figref{f:crossing-edges}).
%(see Figure \ref{fig:crossingArc} (a)).
%so that for each crossing edge,
%one end of the edge lies on a bottom strand while the other
%end lies on a top strand as in Figure \ref{fig:crossingArc} (a).

\begin{figure} 
\input{fig-crossing-edges.tikz}
\label{f:crossing-edges}
\caption{Splitting and flattening crossing edges.}
%\centering 
%\begin{tabular}{cc}
%\includegraphics[width=4cm]{crossingArc.png}
%& \includegraphics[width=5cm]{crossingPush.png}
%\\
%(a)
%& (b)
%\end{tabular}
%\caption{(a) The black strands are
%part of the link and the red strand is the crossing edge.
%(b) The blue and
%red edges represent the split crossing edges and the shaded half disk is bounded
%by the crossing circle} 
%\label{fig:crossingArc} 
\end{figure}

Next we construct the graph that will define the ``top'' torihedron.
We view the link from the point at infinity from the top end
($\torus \times \{1\}$) of the thickened torus.
At each crossing of $K$,
push the top strand towards the bottom strand,
splitting the crossing edge into two
identical edges and spreading them apart as in
\figref{f:crossing-edges}.
%Figure \ref{fig:crossingArc}.
% (b).
%We push the link
%components to infinity and stretch the crossing edge so that we have flattened
%the link onto $\torus \times \{0\}$ except for the crossing circles,
%which remains perpendicular to the projection plane.

@@

\begin{figure}[ht]
\input{cut-slice-flatten.tikz}
\caption{Removing half-twists from augmented twist-regions}
\label{f:cut-slice-flatten}
\end{figure}

@@

\begin{remark}
We note that our main \thmref{t:auglink_hyp}
requires that all non-trivial twist regions be augmented,
but \prpref{p:tori_decomp} does not require it.
\label{r:unnecessary-augment-bigon}
\end{remark}

@@

There exists a non-degenerate
%angle-splitting of $(\Gamma,\theta_\bullet)$
angle-splitting of $(\Gamma,f^*)$
if and only if the following is satisfied:

@@

Recall that $\Gamma_T(L),\Gamma_B(L)$ are
obtained by bow-tie modifications of the diagram $D'$
of a link $K'$
(see \defref{d:tori-decomp-graph}).
Assign to each edge $e$ of $D'$ the angle $\theta_e = \pi/2$
(so that $\theta_e^* = \pi/2$ too).
Using the fact that $K'$ is weakly prime
(which easily follows from $K$ being weakly prime),
it is not hard to see that the condition on cocycles
of \thmref{t:bs-thm4} is satisfied by this assignment.
Thus, there exists a non-degenerate angle-splitting
$\vphi_\bullet$ of $(D',\theta_\bullet^*)$.
%Recall from the proof of \prpref{p:tori_decomp}
%that the two graphs $\Gamma_T(L),\Gamma_B(L)$
%can be obtained by modifying the link diagram of $K$:
%we take $K'$ to be the link that agrees with $K$ everywhere
%except each augmented twist region has only one crossing,
%and $\Gamma_T(L),\Gamma_B(L)$ are

@@

We check that upon gluing the top and bottom torihedra,
the sum of interior dihedral angles $\theta^*$
around each edge is $2\pi$:
crossing edges have $\theta^* = \pi/2$,
and appear four times, twice in each torihedron,
while for bow-tie edges,
simply check for half-twist and non-half-twist cases
separately.

Now we have a decomposition of the two torihedra into
degenerate base-angled pyramids
(recall that that means some of the
interior dihedral angles $\theta^*$ are 0);
since we need the pyramids to be non-degenerate,
we modify the graph on the torihedra
and the angle assignments to make all $\theta^*$ nonzero
as follows.

@@

\begin{figure}
\input{fig-add-edges.tikz}
\caption{Adding edges to non-bow-tie faces;
the $+/-$ labels are short for $+1/-1$.}
%edge indicates an added edge to $\Gamma_T(L)$;
%indicate whether to
%increase/decrease $\vphi$ (and $\theta^*$)
\label{f:adding_edges}
\end{figure}

@@

Now fix some small $\veps > 0$.
Let $\vphi_{\vec{e}}' = \vphi_{\vec{e}} + x \cdot \veps$,
where $x$ is the label on $\vec{e}$ (0 if unlabeled).
Let $\theta_e'^* = \vphi_{\vec{e}}' + \vphi_{\cev{e}}'$.
It is easy to check that the sum of $\theta'^*$
around each vertex is the same as for $\theta^*$,
so it defines an angled torihedron $(\sT_T',\theta'^*)$.
For each face of $\Gamma_T'$, the $+/-$ labels
on the inner side of boundary edges cancel out:
for bow-tie faces, the short edge gets a $+1$ and the long edge
gets $-1$, while for non-bow-tie faces,
it is clear from \figref{f:adding_edges}.
Hence, $\vphi'$ is an angle-splitting of $(\sT_T',\theta'^*)$.
Furthermore, all shorts edges
(which are the only edges with $\vphi_{\vec{e}} = 0$)
have a $+1$ label on each side,
so $\vphi'$ is non-degenerate.

@@

By construction, under the gluing of
$\sT_T(L)$ to $\sT_B(L)$,
the new edges added to $\Gamma_T(L)$
are glued to the new edges added to $\Gamma_B(L)$,
since they are added by the same procedure.
As noted before, upon gluing $\sT_T(L)$ to $\sT_B(L)$
the sum of exterior dihedral angles $\theta^*$ around each edge
is $2\pi$.
This clearly remains true after adding the new edges
(they're label $\pi$ on each torihedron).
Again by construction,
the $+/-$ labels coming from the top and bottom diagrams get canceled out.
Thus, upon gluing $\sT_T'$ to $\sT_B'$,
the sum of new exterior dihedral angles $\theta'^*$ around each edge
is $2\pi$.

@@

%%% comment }

@@

%crossing edge, Menasco thing
%\label{fig:crossingArc} 
%\label{f:crossing-edges}.
